---
author: Aman Das
title: Statistics Undergraduate
format:
  classic-cv-pdf: default
cvcolor: 29aae2 #without `#`
cvfont: cantarell #texlive fonts like raleway,fetamont,gillius,comfortaa
cvmeta:
  - title: Field
    left: Data Science, 
    right: "[aman9das.github.io](https://aman9das.github.io)"
  - title: Techs
    left: R, Quarto, Docker, bash, Python, MySQL
    right: "[amandas62640@gmail.com](mailto:amandas62640@gmail.com)"
  - title: Status
    left: Bachelor of Statistics
    right: "[7550961941](tel:+917550961941)"
cvfooter:
  left: "[Github](https://github.com/Aman9das)"
  right: "[LinkedIn](https://www.linkedin.com/in/aman-das-9a102a1bb/)"
execute:
  echo: false
---

```{r}
#| label: cvevents
cvevents <- function(tbl, when, what, where, details) {

    command_start <- "\\cvevent"

    res <- paste0(
        command_start, "{", tbl[[when]], "}", 
        "{", tbl[[what]], "}",
        "{", tbl[[where]], "}")
    
    tbl[[details]] <- sapply(tbl[[details]], function(x) paste0("{", x, "}", collapse = ","))
    res <- paste0(res, "{",tbl[[details]],"}")

    cat(res, sep = "\n")
}
```

# Summary

Highly motivated statistics student with experience in software design. Interested in various facets of Open Technologies. Puts care into making the analyses easily reproducible. I am at my best when collaborating with others.

# Projects

```{r}
#| label: experience
#| output: asis
tibble::tribble(
    ~role, ~institution, ~dates, ~tasks,
    "Project Lead", "\\href{https://aman9das.github.io/blog/2023-sem2-midsem-stat-project/}{Statistics Project}", "May. 2023",
    c("Coordinate the project to analyse Foreign Tourist Visits to India", "Implemented the analysis paper using Quarto."),
    "Analysis and Design Lead", "\\href{https://aman9das.github.io/blog/2022-sem1-endsem-stat-project/PPT_Reveal.html}{Statistics Project}", "Dec. 2022",
    c("Statistical Analysis of GDP per capita, Sanitation and Life Expectancy across Nations.", "Implemented the presentation design in Quarto."),
    "UI Design", "\\href{https://f-droid.org/en/packages/de.monocles.chat/}{Monocles Chat}", "Sep. 2021",
    c("Redesigned the chat screen.", "Created a new color theme."),
    "Package Deployment", "\\href{https://flathub.org/}{Flathub}", "May 2022 - Present", c("Packaged Remnants of the Precursors (ROTP), Wick Editor, Dino and Gargoyle.", "Maintain Ghostwriter, ROTP and Wick Editor for Flathub."),
    "App Developer", "\\href{https://arcticons.onnno.nl}{Arcticons}", "Oct. 2021", c("Re-implemented the Android app on Candybar framework and created Widgets.", "Developed the UI and added a number of Icons.")
    ) |>
    cvevents(when = "dates", what = "role", where = "institution", details = "tasks")
```

# Education

```{r }
#| label: education
#| output: asis
tibble::tribble(
    ~degree, ~uni, ~dates, ~details,
    "B Stat (Hons)", "Indian Statistical Institute, Kolkata", "2025",
    "First year aggregate: 75\\%",
    "SSCE", "DAV Public School", "2022", c("Aggregate: 88\\%", "KVPY SA: AIR 726"),
    "ICSE", "St. Agnes School", "2020", c("Aggregate: 94\\%", "Proficiency in English, Hindi and Bengali.")
) |>
    cvevents(when = "dates", what = "degree", where = "uni", details = "details")
```

# Interests

\setlength{\tabcolsep}{48pt}
\begin{tabular}{ c c c }
 Distribution Theory & Statistical Inference & Linear Algebra \\
 R & Python & Kotlin \\
 Reproducibility & Design & Open Source
\end{tabular}
